---
title: "Kickstarter - Code"
format: html

---

Data and libraries 

```{r}
#| eval: false
library(readr)
library(lubridate)
library(rpart)
library(rpart.plot)
library(chron)
library(dplyr)
library(stringr)
library(ggplot2)
library(coin)
kickstarter = read_csv("~/documents/kickstarter.csv")

```


Data transformation
```{r}
#| eval: false
kickstarter$category=as.factor(kickstarter$category)
kickstarter$main_category=as.factor(kickstarter$main_category)
kickstarter$currency=as.factor(kickstarter$currency)
kickstarter$state=as.factor(kickstarter$state)
kickstarter$country=as.factor(kickstarter$country)
kickstarter=kickstarter %>% mutate(launch_date=as.Date(launched)) %>% mutate(period_length=deadline-launch_date,month = month(launch_date),day = day(launch_date),launch_season = case_when(
    month > 2 & month < 6 ~ "spring",
    month > 5 & month < 9 ~ "summer",
    month > 8 & month < 12 ~ "autumn",
    month > 11 | month < 3 ~ "winter",
    TRUE ~ "unknown"),launch_weekday=weekdays.Date(kickstarter$launch_date))
kickstarter=kickstarter %>% mutate(launch_weekday=weekdays.Date(kickstarter$launch_date),gen_category=str_trim(paste(main_category,", ",category)),name_length=str_length(name),number_of_words=str_count(name,'\\w+'),n_caps_in_name=str_count(name, "\\b[A-Z]{2,}\\b"),has_caps_in_name=case_when(n_caps_in_name>0 ~ 1, TRUE ~ 0),launch_hour=hour(launched), launch_dayPart = case_when(
    launch_hour > 5 & launch_hour < 13 ~ "morning",
    launch_hour > 11 & launch_hour < 18 ~ "afternoon",
    launch_hour > 17 & launch_hour < 23 ~ "evening",
    launch_hour > 22 | launch_hour < 6 ~ "night",
    TRUE ~ "unknown"
  ),launch_is_weekend=is.weekend(launched),cat_is_cat=case_when(as.character(main_category)==as.character(category)~1,TRUE~0),name_has_marks=as.factor(case_when(str_detect(name,":")==TRUE|str_detect(name,";")==TRUE|str_detect(name,"!")==TRUE~1,TRUE~0)))
                                                                          
kickstarter$launch_season=factor(kickstarter$launch_season,levels=c("autumn","winter","spring","summer"))

kickstarter$launch_weekday=factor(kickstarter$launch_weekday,levels=c("Monday","Tuesday", "Wednesday","Thursday","Friday","Saturday","Sunday"))

kickstarter$has_caps_in_name=as.factor(kickstarter$has_caps_in_name)
kickstarter$cat_is_cat=as.factor(kickstarter$cat_is_cat)
kickstarter$name_has_marks=as.factor(kickstarter$name_has_marks)
kickstarter$gen_category=kickstarter$gen_category %>% str_replace_all(" ,  ",", ")

kickstarter=kickstarter %>% mutate(gen_cat_length=str_length(gen_category),gen_cat_number_words=str_count(gen_category,'\\w+'),gen_cat_first_letter=str_extract(gen_category,'[A-Z]'),cat_length=str_length(category),cat_number_words=str_count(category,'\\w+'),cat_first_letter=str_extract(category,'[A-Z]'))

kickstarter$gen_category=as.factor(kickstarter$gen_category)
kickstarter$launch_dayPart=factor(kickstarter$launch_dayPart,levels=c("morning","afternoon","evening","night"))
kickstarter$gen_cat_first_letter=as.factor(kickstarter$gen_cat_first_letter)
kickstarter$cat_first_letter=as.factor(kickstarter$cat_first_letter)
```




Projects by sum graph
```{r}
#| eval: false
kickstarter_sum_by_category=kickstarter %>% group_by(state,main_category) %>% filter(state=="successful") %>% summarise(average_sum=round(mean(usd_goal_real),1)) %>% arrange(-average_sum) %>% ungroup() %>% dplyr::select(-state)
knitr::kable(kickstarter_sum_by_category,col.names=c("Основная категория","Средняя сумма реализованных проектов"))
```

Success of projects by category
```{r}
#| eval: false
kickstarter_num_cat=kickstarter %>% group_by(main_category) %>% summarise(cnt=n())
kickstarter_success_perc=kickstarter %>% filter(state=="successful") %>% group_by(main_category) %>% summarise(s_cnt=n())
kickstarter_success_by_cat=left_join(kickstarter_success_perc,kickstarter_num_cat,by="main_category") %>% mutate(success_percentage=s_cnt/cnt*100) %>% arrange(-success_percentage) %>% select(-s_cnt,-cnt)
knitr::kable(kickstarter_success_by_cat,col.names=c("Основная категория","Процент успешных проектов"))
```


Graph Success of projects by category
```{r}
#| eval: false
ggplot(data=kickstarter)+geom_bar(aes(x=main_category,fill=state),stat="count")+xlab('Название основной категории')+ylab('Количество проектов')+ggtitle('Успешность проектов в разных категориях')+coord_flip()+scale_fill_brewer(name="Состояние проекта",labels=c("неуспешный","успешный"),palette="Pastel1")
```


Graph Success of projects by Launch hour
```{r}
#| eval: false
ggplot(data=kickstarter)+geom_bar(aes(x=launch_hour,fill=state),position="fill")+ylab("Доля проектов")+xlab("Час запуска")+scale_fill_brewer(name="Состояние проекта",labels=c("неуспешный","успешный"),palette="Pastel1")+ggtitle("Успешность проектов по часу запуска")

```



Graph Success of projects by title length
```{r}
#| eval: false
ggplot(data=kickstarter)+geom_boxplot(aes(x=state,y=name_length, fill=state))+ylab("Длина названия проекта в символах")+xlab("Состояние проекта")+scale_x_discrete(labels=c("Неуспешный", "Успешный"))+ggtitle("Успешность проекта по длине названия проекта")+scale_fill_brewer(name="Состояние проекта",labels=c("неуспешный","успешный"),palette="Pastel1")
```



Graph Success of projects by special symbols in the title
```{r}
#| eval: false
ggplot(data=kickstarter)+geom_bar(aes(x=name_has_marks,fill=state),position="fill")+xlab("Наличие символов :,!,; в названии проекта")+ylab("Доля проектов")+ggtitle("Успешность проекта по наличию символов в названии")+
scale_fill_brewer(name="Состояние проекта",labels=c("неуспешный","успешный"),palette="Pastel1")+scale_x_discrete(labels=c("Нет", "Да"))
```



Model creation
```{r}
#| eval: false
# "seed" for the random number generator
set.seed(1234) 

# Let's take 80% as training
kickstarter_train = kickstarter %>% dplyr::sample_frac(.8)

# create a test dataset
# using an anti-join to remove all observations that were in the training set
kickstarter_test = dplyr::anti_join(kickstarter, kickstarter_train, by = 'id') %>% dplyr::select(-id)

kickstarter_train = kickstarter_train %>% dplyr::select(-id)

## Building a Complex Model
## Let's try building a complex model that will give us the highest accuracy. We'll use the rpart package and all variables except id, name, date, deadline, launched, backers, and launch date.

### Unpruned Tree
tree1 <- rpart(state ~ category+gen_category+main_category+usd_goal_real+day+country+month+launch_season+currency+period_length+launch_weekday+name_length+number_of_words+has_caps_in_name+n_caps_in_name+launch_dayPart+launch_hour+launch_is_weekend+cat_is_cat+name_has_marks+cat_length+cat_number_words+cat_first_letter+gen_cat_length+gen_cat_number_words+gen_cat_first_letter, method= "class",data = kickstarter_train, control = rpart.control(cp = 0.001)) 
rpart.plot::rpart.plot(tree1)

#predict
pred.kick_train1 = predict(tree1, type="class")

## **accuracy on the training set**
t1 = table(pred.kick_train1, kickstarter_train$state)
acc1tr=(t1[1,1] + t1[2,2])/sum(t1)
acc1tr

## **accuracy on the test set**
pred.kick_test1 = predict(tree1, newdata = kickstarter_test, type="class")
t11 = table(pred.kick_test1, kickstarter_test$state)
acc1ts=(t11[1,1] + t11[2,2])/sum(t11)
acc1ts

## As we can see, the accuracy on the test sample was lower, but only slightly. Let's prune the tree based on the optimal cp value.

## **Building the cp table**
set.seed(1234)
cp_tbl = printcp(tree1)
set.seed(1234)
min_cp = cp_tbl %>%
  as.data.frame() %>% 
  arrange(xerror) %>% 
  head(n = 1)
min_cp

## **prune the tree using prune()**
tree2 = prune(tree1, cp = 0.00102)

rpart.plot::rpart.plot(tree2,fallen.leaves = FALSE,faclen=2)
# making a prediction
pred.kick_train2=predict(tree2, type="class")

## **accuracy for training sample**
t2 = table(pred.kick_train2, kickstarter_train$state)
acc2tr=(t2[1,1] + t2[2,2])/sum(t2)
acc2tr

## **accuracy for the test sample**
#prediction
pred.kick_test2 = predict(tree2, kickstarter_test, type="class")
#accuracy
t22 = table(pred.kick_test2, kickstarter_test$state)
acc2ts=(t22[1,1] + t22[2,2])/sum(t22)
acc2ts
## So, we've built a complex model and achieved an accuracy of 0.7018865 on the training set and 0.6823584 on the test set. Let's try building a less complex model that still achieves similar accuracy.
## To do this, use the summary command to see which variables Model 1 considered most important and build a model based on them.

## Less complicated model creation

## **model and accuracy for the training set**


#summary(tree2)
tree0 <- rpart(state ~ gen_category+category+main_category+usd_goal_real+number_of_words+cat_is_cat+launch_dayPart+period_length+cat_length+cat_number_words+cat_first_letter+gen_cat_length+gen_cat_number_words+gen_cat_first_letter+country+currency+launch_weekday+launch_hour, method="class",data=kickstarter_train)
pred0 = predict(tree0, type="class")
#function for line breaks for variable names
split.fun <- function(x, labs, digits, varlen, faclen)
{
    # replace commas with spaces (needed for strwrap)
    labs <- gsub(",", " ", labs)
    for(i in 1:length(labs)) {
        # split labs[i] into multiple lines
        labs[i] <- paste(strwrap(labs[i], width=32), collapse="\n")
    }
    labs
}
prp(tree0, split.fun=split.fun,fallen.leaves=FALSE,extra=4,cex=0.45)

t0 = table(pred0, kickstarter_train$state)

(t0[1,1] + t0[2,2])/sum(t0)

## **accuracy for test sample**
pred00 = predict(tree0, kickstarter_test, type="class")

t = table(pred00, kickstarter_test$state)

(t[1,1] + t[2,2])/sum(t)
```



Итак, мы построили намного более простую модель, имеющую точность 0.6758971 для тренировочной и 0.6693204 для тестовой выборок. 

```{r}
a=(0.7018865-0.6758971)*100
b=(0.6823584-0.6693204)*100
paste("At the same time, the accuracy of the simplified model is ",round(a,2),"% (training sample) and ",round(b,2),"% (test sample) lower than that of an order of magnitude more complex model")
```

Importance of variables in the model
```{r}
treecat = c('category','gen_category','main_category','cat_first_letter','gen_cat_first_letter','cat_length','usd_goal_real','period_length','name_length','country','launch_dayPart','number_of_words','currency','launch_hour')
meaning = c('category', 'project category and subcategory', 'main category', 'first letter of category name', 'first letter of subcategory name', 'category name length in characters', 'project target amount in dollars', 'fundraising duration', 'project name length in characters', 'country', 'part of the day the project was launched', 'number of words in the project name', 'currency', 'launch hour')
importance=c('25','22','12','11','10','7','4','3','1','1','1','1','1','1')
catable = data.frame(treecat, meaning,importance)
knitr::kable(catable, col.names = c("Variable","Name Explanation","Significance"))
```

